---
title: "ESM 204: Homework #1"
author: "Natalie Smith"
format: 
  html:
    code-fold: true
    toc: true
    number-sections: true
    embed-resources: true
theme: Lux
editor: visual
execute:
  echo: true
  message: false
  warning: false
---
### Basics of R:

a) Launch RStudio and start a new R script (or markdown document, up to you!). Write a simple program that counts from 1 to 10 by increments of 0.5. Print out the results.
```{r}
# Count from 1 to 10 by increments of 0.5
count <- seq(from = 1, to = 10, by = 0.5)

# Print out the results
print(count)
```
b) Write a custom function. Your function should have three inputs (X, a, and b) and one
output (Y). The function should calculate Y = aX + b. Once you have created it, call your function for the values X=5, b=10, and a=3. What value of Y did you get? Confirm algebraically that it is correct.

```{r}
# write a custom function with three inputs (X,a,b) and one output (Y). The function should calculate Y = aX + b.
custom_function <- function(X, a, b){
  Y <- a*X + b
  return(Y)
}

# call the function for the values X=5, b=10, and a=3
custom_function(5, 3, 10)

```

